---
title: "Assignment 4: Exploratory Data Analysis"
author: "Nikkos Miyasato"
date: "`r Sys.Date()`"
format: html
editor: visual
---

## Overview

Due: **Friday 9/25 11:59pm**

In this assignment, you will practice Exploratory Data Analysis (EDA) by making plots from a **Kaggle dataset of your choice**.

You will submit **at least 4 different plot types**, with short written interpretations for each.

Part 1 is worth a total of 1 point, and Part 2 is worth 4 points (1 point for each plot, question, and interpretation).

## Part 1: Get Your Data

1.  Go to [Kaggle Datasets](https://www.kaggle.com/datasets), create an account, and choose a dataset that interests you. Ensure your data has:
    1.  At least 4 variables
    2.  At least one categorical and one continuous variable
    3.  I recommend data that contains at least 50-100 observations
    4.  Tip: Open the CSV on Kaggle first and skim the columns to check that it meets these requirements
2.  Download the dataset as a CSV file.
3.  Place the CSV in your class R project `data` folder.
4.  Read it into R with `read_csv()`.

```{r, message=FALSE}
#load necessary packages
library(tidyverse)
library(janitor)
library(ggplot2)
library(plotly)
library(dplyr)

#read dataframe
GlobalSecurityThreats <- read_csv("data/GST_2015-2024.csv")

#clean names so data is easier to work with
GlobalSecurityThreats <- GlobalSecurityThreats |> 
  clean_names()

```

### Describe your data

**Data author:** Atharva Soundankar

**Link:** https://www.kaggle.com/datasets/atharvasoundankar/global-cybersecurity-threats-2015-2024

**Description:** The Global Cyber security Threats Data set (2015-2024) provides in depth exploration on various cyber attacks, types of Malware's used, what industries were targeted, and the affected countries of the attacks. This data set is used for threat intelligence analysis and to assist in AI Machine Learning development to enhance global security in the field of cyber intrusion.

### Preview your data

Use some of the strategies you've learned in this class or picked up elsewhere to summarize and quickly explore your data.

```{r}
#view the data in a seperate tab
view(GlobalSecurityThreats)

```

### Summary stats

Before you make plots, you must also:

- Show basic summaries of your dataset, including:

  - `count()` for categorical variables (to see category frequencies)

  - `mean()`, `median()`, `sd()`, `min()`, `max()` for at least one numeric variable

  - Remember to use `na.rm = TRUE` if you have missing values that affect summaries!

```{r}

#show a basic summary of dataframe
summary(GlobalSecurityThreats)

#Count frequncy of variables
GlobalSecurityThreats |>  
  count("attack_source")
GlobalSecurityThreats |> 
  count("attack_type")

#show head of data - 10 rows
head(GlobalSecurityThreats, n = 10)

#mean of financial loss in $
mean(GlobalSecurityThreats$"financial_loss_in_million", na.rn = TRUE)

#median of finaincial loss in $
median(GlobalSecurityThreats$"financial_loss_in_million", na.rn = TRUE)

#Standard Deviation from loss in $
sd(GlobalSecurityThreats$"financial_loss_in_million")

#Minimum loss of $
min(GlobalSecurityThreats$"financial_loss_in_million", na.rn = TRUE)

#maximim loss of $
max(GlobalSecurityThreats$"financial_loss_in_million", na.rn = TRUE)
```

Write 2–3 sentences here: What do the category counts tell you? What does the numeric summary show about spread, center, or outliers?

**ANSWER:** In count, I've chosen to pull data from Attack Source (amount of attacks via sources) and Attack Type (amount of specified attacks). I've noticed that despite removing the na variables there is still a slight variance from Mean and Median - 50.49297 vs 50.795. There is also a high Standard Deviation in loss of \$ by 28 million.

## Part 2: Visualize Your Data

Create **at least 4 plots**, each of a different type. Your plots may share variables, but each one should answer a **different question** about your data. At least one plot must include a variable that isn't in the others.

Each plot should have:

- A title and labeled axes

- A short interpretation (2-3 sentences) underneath

  - Sample interpretation: (boxplot of `price` by `cut` from `diamonds` data): *Ideal diamonds have the lowest median price, which is surprising since Ideal is the best cut. Every cut also has a lot of very expensive outliers. One possible reason for the low Ideal prices is that Ideal diamonds might be smaller on average, which I could check by plotting carat by cut.*

Use the EDA Part 3 notes as your reference. Since this is an EDA assignment, your plots can look plain, but consider adding colors, themes, or interactivity for interpretability and visual appeal if you know how.

### Plot 1

```{r}
country_attacked <- ggplot(data = GlobalSecurityThreats, 
                           aes(x = number_of_affected_users, y = country,
                               color = incident_resolution_time_in_hours)) +
  geom_point() +
  theme_bw() +
  labs(title = "Scatter Plot of Country Attacked",
       x = "Number of Attacks",
       y = "Country's Attacked",
       color = "Time in Hours")

ggplotly(country_attacked)
```

**Question this plot answers:** What is the frequency of attacks a country received from the report, and how long was the attack active?

Interpretation: There is variance in the time each attack had lapsed within each country. They range from approximately 10 hours in length to approximately 60 hours in length. The majority of attacks when scrutinized appear to be commonplace in the 0-40 range

### Plot 2

```{r}
boxgraph <- ggplot(data = GlobalSecurityThreats, 
       aes (x = year, y = target_industry, fill = country))+
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  labs(title = "Boxplot of Target Industry",
       x-axis = "Target Industry",
       y-axis = )

ggplotly(boxgraph)
```

**Question this plot answers:** In the years if this study what were the target industries that were affected by each country?

Interpretation:

### Plot 3

```{r}
# insert code here

```

Question this plot answers:

Interpretation:

### Plot 4

```{r}
# insert code here

```

Question this plot answers:

Interpretation:

------------------------------------------------------------------------

# Push to GitHub AND submit link on Canvas!!

Assignment is due Friday 9/25 at 11:59 pm.
