---
title: "CS-201 First Look at R"
author: "Nikkos Miyasato"
date: 8/26/26
format: html
editor: visual
toc: true
---

This is the first class session walking through code together!

We will be programming in R and in this class will primarily use `qmd` documents to keep things tidy and presentable. Each code chunk below is where you will run R code. Feel free to add notes as we go along.

## Navigating RStudio

By default, everyone should have the same four-pane layout:

1.  **Top left**: Source editor

2.  **Bottom left**: Console (and Terminal and Background Jobs)

3.  **Top Right**: Environment

4.  **Bottom Right**: Files, Plots, Help, etc.

### Customizing look & feel

Feel free to customize your RStudio.\
Go to Tools \> Global Options \> Appearance to select a color scheme you like.

I also personally like to enable these:

- Tools \> Global Options \> Code \> Insert matching parens/quotes

- Tools \> Global Options \> Console \> Show syntax highlighting in console input

I don't recommend rearranging your pane layout at this time.

## Running Code

Try running some code

Load Library

```{r}

library(tidyverse)
```

Some scripting

```{r}
#This is how to assign value to name
quiz1 <- 88
quiz2 <- 94
quiz3 <- 79

#Averageing
(quiz1 + quiz2 + quiz3) / 3

#ex
mean(quiz1 + quiz2 + quiz3)
mean(quiz1, quiz2, quiz3)

#in mean EX will take first syntax argument
mean(c(quiz1, quiz2, quiz3))

list <- (c(quiz1, quiz2, quiz3))

average <- mean(list)

quiz3 <- 99
quiz3

list <- (c(quiz1, quiz2, quiz3))
average <- mean(list)
average
round(average, 2)
```

More scores

```{r}
scores <- c(88, 94, 78, 91, 85)
scores

mean(scores)
min(scores)
max(scores)
list(scores)

summary(scores)

students <- c("a", "b", "c", "d", "e")
students

# Bar plot
barplot(scores, names.arg = students)

## More detailed/organized barplot
barplot(scores, names.arg = students, ylim = c(0, 100), main = "TITLE", xlab = "PUPILS", y = "what they got")

```

## Introducing `ggplot2`

#### \*Make a Data Frame

```{r}

grades_df <- data.frame(student = students, score = scores)

grades_df

summary(grades_df)

nrow(grades_df)
#num of rows

ncol(grades_df)
#num of columns

dim(grades_df)
#num of dimesions
```

Try GGPLOT

```{r}

ggplot(data = grades_df, aes(x = student, y = score))
#foundation that creates a graph, Layer1

ggplot(data = grades_df, aes(x = student, y = score)) + geom_col()
#2nd layer for detail

ggplot(data = grades_df, aes(x = student, y = score)) + geom_col(fill = "steelblue")
#Third layer to change color

ggplot(data = grades_df, aes(x = student, y = score)) + geom_col(fill = "steelblue") + geom_point()
#Adds points

ggplot(data = grades_df, aes(x = student, y = score)) + geom_col(fill = "steelblue") + geom_point() + geom_line()

ggplot(data = grades_df, aes(x = student, y = score)) + geom_col(fill = "steelblue") + geom_point() + geom_line() + theme_minimal()

ggplot(data = grades_df, aes(x = student, y = score)) + 
  geom_col(fill = "steelblue")+ 
  geom_point() + 
  geom_line() + 
  theme_minimal() + 
  labs(title = "BEST GRAPH MADE",#to create labels
         subtitle = "some quiz scores",
         x = "the students",
         y = "what did they score on their quiz")

ggplot(data = grades_df, aes(x = student, y = score)) + 
  geom_col(aes(fill = student)) +
  scale_fill_manual(values = c(
    "a" = "red",
    "b" = "blue",
    "c" = "green",
    "d" = "purple",
    "e" = "orange"
  )) +
  theme_minimal()
#This is how you can add color depth by assigning color to individual values
```

Mutate the dataframe

```{r}

library(tidyverse)

name <- grades_df |>
  mutate(study_hours = c(2, 5, 1, 4, 9))
#pipe operator - ctrl + SHIFT + m

grades_df <- grades_df |>
  mutate(study_hours = c(2, 5, 1, 4, 9))
#
```

Make a scatterplot

```{r}
library(ggplot2)

ggplot(data = grades_df, aes(x = study_hours, y =score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

#Cannot use fill (use color instad), don't use reorder
```

## Extra Credit (1pt)

What result do you expect when you run this expression?

```{r}

1 + 1 + 1

quiz1 + quiz2

1 + 1 + 1 - 3

0.1 + 0.1 + 0.1 - 0.3
#Assignment is find find out why
## The reason this code did not output zero is because RStudio code is binary and not by decimal value. R recognized "0.1" and "0.3" as floating-point approximate numbers - thus returning the value of 5.551115e-17. This value is a mathmatical expression that represents 5.551115×10−17

options(scripen = 999)

print(0.1, 20)
print(0.3, 20)
print (0.5, 20)

#{python} - can inject python into R-Chunk
```

What did you observe?

> Answer: When the code was ran the return product was:
>
> ```         
> 5.551115e-17
> ```

Why do you think this is happening? (You can look online for help)

> Answer: The reason this code did not output zero is because R code is operating on binary values and not by decimal value. R recognized "0.1" and "0.3" as floating-point approximate numbers - thus returning the value of 5.551115e-17. This value is a mathmatical expression that represents 5.551115×10−17 and gives a rounding error.
