---
title: "CS-201 First Look at R"
author: "Nikkos Miyasato"
date: 8/26/26
format: html
editor: visual
toc: true
---

This is the first class session walking through code together!

We will be programming in R and in this class will primarily use `qmd` documents to keep things tidy and presentable. Each code chunk below is where you will run R code. Feel free to add notes as we go along.

## Navigating RStudio

By default, everyone should have the same four-pane layout:

1.  **Top left**: Source editor

2.  **Bottom left**: Console (and Terminal and Background Jobs)

3.  **Top Right**: Environment

4.  **Bottom Right**: Files, Plots, Help, etc.

### Customizing look & feel

Feel free to customize your RStudio.\
Go to Tools \> Global Options \> Appearance to select a color scheme you like.

I also personally like to enable these:

- Tools \> Global Options \> Code \> Insert matching parens/quotes

- Tools \> Global Options \> Console \> Show syntax highlighting in console input

I don't recommend rearranging your pane layout at this time.

## Running Code

Try running some code

```{r}


```

## Introducing `ggplot2`

```{r}

```

## Extra Credit (1pt)

What result do you expect when you run this expression?

```{r}

```

What did you observe?

> Answer:

Why do you think this is happening? (You can look online for help)

> Answer:
