---
title: "CS-201 First Look at R"
author: "Nikkos Miyasato"
date: 8/26/26
format: html
editor: visual
toc: true
---

This is the first class session walking through code together!

We will be programming in R and in this class will primarily use `qmd` documents to keep things tidy and presentable. Each code chunk below is where you will run R code. Feel free to add notes as we go along.

## Navigating RStudio

By default, everyone should have the same four-pane layout:

1.  **Top left**: Source editor

2.  **Bottom left**: Console (and Terminal and Background Jobs)

3.  **Top Right**: Environment

4.  **Bottom Right**: Files, Plots, Help, etc.

### Customizing look & feel

Feel free to customize your RStudio.\
Go to Tools \> Global Options \> Appearance to select a color scheme you like.

I also personally like to enable these:

- Tools \> Global Options \> Code \> Insert matching parens/quotes

- Tools \> Global Options \> Console \> Show syntax highlighting in console input

I don't recommend rearranging your pane layout at this time.

## Running Code

Try running some code

Load Library

```{r}

library(tidyverse)
```

Some scripting

```{r}
#This is how to assign value to name
quiz1 <- 88
quiz2 <- 94
quiz3 <- 79

#Averageing
(quiz1 + quiz2 + quiz3) / 3

#ex
mean(quiz1 + quiz2 + quiz3)
mean(quiz1, quiz2, quiz3)

#in mean EX will take first syntax argument
mean(c(quiz1, quiz2, quiz3))

list <- (c(quiz1, quiz2, quiz3))

average <- mean(list)

quiz3 <- 99
quiz3

list <- (c(quiz1, quiz2, quiz3))
average <- mean(list)
average
round(average, 2)
```

More scores

```{r}
scores <- c(88, 94, 78, 91, 85)
scores

mean(scores)
min(scores)
max(scores)
list(scores)

summary(scores)

students <- c("a", "b", "c", "d", "e")
students

# Bar plot
barplot(scores, names.arg = students)
```

## Introducing `ggplot2`

```{r}
library(ggplot2)
```

## Extra Credit (1pt)

What result do you expect when you run this expression?

```{r}

```

What did you observe?

> Answer:

Why do you think this is happening? (You can look online for help)

> Answer:
